---
title: "Day 1 Case Study Exercises"
output: github_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE
  )
```

# Prerequisites

All of the exercises can be solved using the `tidyverse` and `completejourney` packages. The `completejourney` package is an R data package that has been created so the full suite of Complete Journey datasets can be loaded as a library. You can find details about the data and the variables at http://bit.ly/completejourney. In order to use the data you must first install the package following these steps:

```{r}
install.packages('completejourney')
```

Go ahead and load the `tidyverse` and `completejourney` packages:

```{r}
library(_______)
library(_______)
```

The exercises that follow will use various data sets included in the `completejourney` package to include:

```{r datasets}
transactions <- transactions_sample  # just using a sample of the entire data

transactions
products
```


# Data Transformation

The following five exercises are based on concepts covered in the data transformation (`dplyr`) slides. 

---

## Exercise 1

Create three new variables named `regular_price`, `loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```


This question is designed to strengthen your ability to use the `dplyr` verb `mutate()` to create new variables from existing ones. It should also help you develop a better understanding of the discount variables in `transactions`.

```{r}
transactions <- transactions %>% 
  mutate(
    regular_price = _______,
    loyalty_price = _______,
    coupon_price = _______
    )
```

---

## Exercise 2

Create three new variables named `regular_price`, `loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```


This question is designed to strengthen your ability to use the `dplyr` verb `mutate()` to create new variables from existing ones. It should also help you develop a better understanding of the discount variables in `transactions`.

```{r}
transactions <- transactions %>% 
  mutate(
    ______,
    ______,
    ______
    )
```

---

## Exercise 3

`transactions` includes 20,902 unique product IDs. How many of these products (not transactions!) had a regular price of one dollar or less? What does this count equal when loyalty price is one dollar or less?  How about when coupon price is one dollar or less?

**Hint:** After filtering, select the `product_id` column and count unique products using the `n_distinct()` function. 

This question is designed to strengthen your ability to use the `dplyr` verbs `filter()` and `select()`.

```{r}
# change your input in filter to assess the different options:
#   - regular_price <= 1
#   - loyalty_price <= 1
#   - coupon_price <= 1

# how many products had a regular price of $1 or less
transactions %>% 
  filter(_______) %>% 
  select(product_id) %>% 
  n_distinct()

# how many products had a loyalty price of $1 or less
transactions %>% 
  filter(_______) %>% 
  select(product_id) %>% 
  n_distinct()

# how many products had a coupon price of $1 or less
transactions %>% 
  filter(_______) %>% 
  select(product_id) %>% 
  n_distinct()

```

---

## Exercise 3

What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use `group_by()` and `summarize()`. Depending on your approach you may or may not use `mutate()`. 

This question is designed to strengthen your ability to use the `dplyr` verbs `group_by()`, and `summarize()`.

```{r}
transactions%>%
  group_by(_______) %>%
  summarize(basket_value = _______) %>%
  summarize(proportion_over_10 = _______)
```

---

## Exercise 4

Which stores had the largest total `sales_value` (hint: `sum(sales_value, na.rm = TRUE)`? Which stores had the largest average loyalty discount as defined below?  

**Hint:** You can calculate loyalty discount as a percentage of regular price using the following logic:

```{r eval = FALSE}
pct_loyalty_disc = 1 - (loyalty_price / regular_price)
```

This question is designed to strengthen your ability to use the `dplyr` verbs `mutate()`, `group_by()`, `summarize()`, and `arrange()`.

```{r}
# find the stores with the largest total `sales_value`
transactions %>%
  group_by(store_id) %>%
  summarize(total_sales_value = _______) %>%
  arrange(_______)
```

```{r}
# find the stores with the largest average (mean) loyalty discount
transactions %>%
  mutate(pct_loyalty_disc = _______) %>%
  group_by(_______) %>%
  summarize(avg_pct_loyalty_disc = _______) %>%
  arrange(_______)
```

---

# Data Visualization

The following five questions are based on concepts covered in the data visualization (`ggplot2`) slides. They can be answered using the `transactions` and `products` datasets.  

---

## Exercise 5

Create a histogram of the `quantity` variable in the `transactions` data. What, if anything, do you find unusual about this visualization? 

This question is designed to strengthen your ability to use `geom_histogram()`.


```{r}
ggplot(data = ______, aes(x = ______)) + 
  geom_histogram()
```

---

## Exercise 6

Use a line graph to plot total sales value by `date` (I have included a mutate statement that parses the YYY-MM-DD from the `transaction_timestamp` variable and names it `date`). What kind of patterns do you find?

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_line()`.

```{r}
transactions %>% 
  mutate(date = lubridate::as_date(transaction_timestamp)) %>%
  group_by(______) %>% 
  summarize(total_sales_value = ______) %>%
  ggplot(aes(x = ______, y = ______)) + 
  geom_line()
```

---

## Exercise 7

Use a bar graph to compare the total sales values of national and private-label brands. 

**Hint**: Because `transactions` does not contain product metadata, run the code below to create a new dataset with additional product information in it. Use `my_transaction_data` for your answer.

```{r}
my_transaction_data <- left_join(transactions, products, by = 'product_id')
```

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_col()`.

```{r}
my_transaction_data %>%
  group_by(brand) %>%
  summarize(total_sales_value = ______) %>%
  ggplot(aes(x = ______, y = ______)) + 
  geom_col()
```

---

## Exercise 8

Building on the previous exercise, suppose you want to understand whether the retailer's customers' preference for national brands (compared to private-label brands) is stronger in the soft drink category than it is in the cheese category. Examine this supposition by using a stacked bar graph to compare the split between national and private-label brands for soft drinks and cheeses. 

**Hint**: Follow these three steps to create your plot: 

 - Filter `my_transaction_data` to include only transactions with `product_category` equal to "SOFT DRINKS" or "CHEESE" 
 - Calculate total sales value by `product_category` and `brand`
 - Create the bars using `geom_col` and include `fill = brand` within `aes()`

```{r}
my_transaction_data %>%
  filter(product_category %in% c('SOFT DRINKS', 'CHEESE')) %>%
  group_by(brand, product_category) %>%
  summarize(total_sales_value = sum(sales_value)) %>%
  ggplot(aes(x = product_category, y = total_sales_value, fill = brand)) + 
  geom_col()
```

---

## Exercise 9

The code below filters `my_transaction_data` to include only peanut better, jelly, and jam transactions. Then it creates a new variable named `package_size` equal to product size in ounces. Create a bar graph with `pb_and_j_data` to visualize the distribution of the retailer's PB\&J transactions by product size. Which two product sizes are the most popular?

```{r Q5, warning = FALSE}
pb_and_j_data <- my_transaction_data %>% 
  filter(product_category == 'PNT BTR/JELLY/JAMS') %>%
  mutate(
    product_size = as.factor(as.integer(gsub('([0-9]+)([[:space:]]*OZ)', '\\1',
                                             package_size)))
  )
```

```{r}
ggplot(pb_and_j_data, aes(x = ______)) + 
  ______
```
