---
title: "Day 1 Case Study Exercises"
output: github_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE
  )
```

# Prerequisites

All of the exercises can be solved using the `tidyverse` and `completejourney` packages. The `completejourney` package is an R data package that has been created so the full suite of Complete Journey datasets can be loaded as a library. In order to use the data you must first install the package following these steps:

```{r, eval=FALSE}
devtools::install_github('bradleyboehmke/completejourney')
```

__NOTE__: Installing packages from GitHub requires the installation of the devtools package, which can be installed by running the following command from the R console: `install.packages('devtools')`.

Go ahead and load the `tidyverse` and `completejourney` packages:

```{r, eval=FALSE}
library(_______)
library(_______)
```

The exercises that follow will use various data sets included in the `completejourney` package to include:

```{r datasets, eval=FALSE}
transactions
products
```


# Data Transformation

The following five questions are based on concepts covered in the data transformation (`dplyr`) slides. Answer them using `transactions` in the Complete Journey data package.

```{r load-packages}
transactions <- transactions %>% 
  select(
    quantity,
    sales_value, 
    retail_disc, coupon_disc, coupon_match_disc,
    household_id, store_id, basket_id, product_id, 
    week_no, day, trans_time
  )
```

---

**Question 1**: Change the discount variables (i.e., `retail_disc`, `coupon_disc`, `coupon_match_disc`) from negative to positive. 

**Hint:** Use the `abs()` function within `mutate()`. 

This question is designed to strengthen your ability to use the `dplyr` verb `mutate()` to overwrite an existing variable.

---

**Question 2**: Create three new variables named `regular_price`, `loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```


This question is designed to strengthen your ability to use the `dplyr` verb `mutate()` to create new variables from existing ones. It should also help you develop a better understanding of the discount variables in `transactions`.

---

**Question 3**: `transactions` includes 92,339 unique product IDs. How many of these products (not transactions!) had a regular price of one dollar or less? What does this count equal when loyalty price is one dollar or less?  How about when coupon price is one dollar or less?

**Hint:** After filtering, select the `product_id` column and count unique products using the `n_distinct()` function. 

This question is designed to strengthen your ability to use the `dplyr` verbs `filter()` and `select()`.

---

**Question 4**: What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use `group_by()` and `summarize()`. Depending on your approach you may or may not use `mutate()`. 

This question is designed to strengthen your ability to use the `dplyr` verbs `group_by()`, `summarize()` and `ungroup()`.

---

**Question 5**: Which store with over $10K in total `sales_value`, discounts its products the most for loyal customers? 

**Hint:** You can calculate loyalty discount as a percentage of regular price using the following logic:

```{r eval = FALSE}
pct_loyalty_disc = 1 - (loyalty_price / regular_price)
```

This question is designed to strengthen your ability to use the `dplyr` verbs `filter()`, `mutate()`, `group_by()`, 
`summarize()`, and `arrange()`.

---

# Data Visualization

The following five questions are based on concepts covered in the data visualization (`ggplot2`) slides. They can be answered using the `transactions` and `products` datasets from the `completejourney` package.  

---

**Question 1**: Create a histogram of the `quantity` variable in the `transactions` data. What, if anything, do you find unusual about this visualization? 

This question is designed to strengthen your ability to use `geom_histogram()`.

---

**Question 2**: Use a line graph to plot total sales value by day. What, if anything, do you find unusual about this visualization?

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_line()`.


---

**Question 3**: Use a bar graph to compare the total sales values of national and private-label brands. 

**Hint**: Because `transactions` does not contain product metadata, run the code below to create a new dataset with additional product information in it. Use `my_transaction_data` for your answer.

```{r}
my_transaction_data <- left_join(transactions, products, by = 'product_id')
```

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_col()`.

---

**Question 4**: Building on Question 3, suppose you want to understand whether the retailer's customers' preference for national brands (compared to private-label brands) is stronger in the soft drink category than it is in the cheese category. Examine this supposition by using a stacked bar graph to compare the split between national and private-label brands for soft drinks and cheeses. 

**Hint**: Follow these three steps to create your plot: 

 - Filter `my_transaction_data` to include only transactions with `commodity_desc` equal to "SOFT DRINKS" or "CHEESE" 
 - Calculate total sales value by `commodity_desc` and `brand`
 - Create the bars using `geom_col` and include `fill = brand` within `aes()`


---

**Question 5**: The code below filters `my_transaction_data` to include only peanut better, jelly, and jam transactions. Then it creates a new variable named `product_size` equal to product size in ounces. Create a bar graph with `pb_and_j_data` to visualize the distribution of the retailer's PB\&J transactions by product size. Which two product sizes are the most popular?

```{r Q5, warning = FALSE}
pb_and_j_data <- my_transaction_data %>% 
  filter(commodity_desc == 'PNT BTR/JELLY/JAMS') %>%
  mutate(
    product_size = as.factor(as.integer(gsub('([0-9]+)([[:space:]]*OZ)', '\\1',
                                             curr_size_of_product)))
  )
```


