---
title: "Day 1 Case Study Solutions"
output: github_document
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE
  )
```

# Prerequisites

All of the exercises can be solved using the `tidyverse` and `completejourney` packages. The `completejourney` package is an R data package that has been created so the full suite of Complete Journey datasets can be loaded as a library. You can find details about the data and the variables at http://bit.ly/completejourney. In order to use the data you must first install the package following these steps:

```{r, eval=FALSE}
install.packages('completejourney')
```

Go ahead and load the `tidyverse` and `completejourney` packages:

```{r}
library(tidyverse)
library(completejourney)
```

The exercises that follow will use various data sets included in the `completejourney` package to include:

```{r datasets}
transactions <- transactions_sample  # just using a sample of the entire data

transactions
products
```


# Data Transformation

The following five exercises are based on concepts covered in the data transformation (`dplyr`) slides. 

---

## Exercise 1

Create three new variables named `regular_price`, `loyalty_price`, and `coupon_price` according to the following logic:

```{r eval = FALSE}
regular_price = (sales_value + retail_disc + coupon_match_disc) / quantity
loyalty_price = (sales_value + coupon_match_disc) / quantity
coupon_price  = (sales_value - coupon_disc) / quantity
```


This question is designed to strengthen your ability to use the `dplyr` verb `mutate()` to create new variables from existing ones. It should also help you develop a better understanding of the discount variables in `transactions`.

#### Answer

```{r}
(transactions <- transactions %>% 
  mutate(
    regular_price     = (sales_value + retail_disc + coupon_match_disc) / quantity,
    loyalty_price     = (sales_value + coupon_match_disc) / quantity,
    coupon_price      = (sales_value - coupon_disc) / quantity
  ) %>% 
  select(regular_price, loyalty_price, coupon_price, everything())
)
```

---

## Exercise 2

`transactions` includes 20,902 unique product IDs. How many of these products (not transactions!) had a regular price of one dollar or less? What does this count equal when loyalty price is one dollar or less?  How about when coupon price is one dollar or less?

**Hint:** After filtering, select the `product_id` column and count unique products using the `n_distinct()` function. 

This question is designed to strengthen your ability to use the `dplyr` verbs `filter()` and `select()`.

#### Answer

2,748 products had a regular price less than or equal to $1.00. This count for loyalty price is 4,648 products. For coupon price, it's 4,844. These numbers indicate that over 2,000 products were discounted to a dollar or less with loyalty-card promotions or coupons.

```{r}
# how many products had a regular price of $1 or less
transactions %>% 
  filter(regular_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

# how many products had a loyalty price of $1 or less
transactions %>% 
  filter(loyalty_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()

# how many products had a coupon price of $1 or less
transactions %>% 
  filter(coupon_price <= 1) %>% 
  select(product_id) %>% 
  n_distinct()
```

---

## Exercise 3

What proportion of baskets are over $10 in sales value? 

**Hint:** You need to use `group_by()` and `summarize()`. Depending on your approach you may or may not use `mutate()`. 

This question is designed to strengthen your ability to use the `dplyr` verbs `group_by()`, and `summarize()`.

#### Answer

Approximately 11% of baskets bring the retailer over $10 in value.

```{r}
transactions%>%
  group_by(basket_id) %>%
  summarize(basket_value = sum(sales_value)) %>%
  summarize(proportion_over_10 = mean(basket_value > 10))
```

---

## Exercise 4

Which stores had the largest total `sales_value` (hint: `sum(sales_value, na.rm = TRUE)`? Which stores had the largest average loyalty discount as defined below?  

**Hint:** You can calculate loyalty discount as a percentage of regular price using the following logic:

```{r eval = FALSE}
pct_loyalty_disc = 1 - (loyalty_price / regular_price)
```

This question is designed to strengthen your ability to use the `dplyr` verbs `mutate()`, `group_by()`, `summarize()`, and `arrange()`.

#### Answer

`store_id == 367` has has the largest total sales value and `store_id == 224` has
the largest average loyalty discount percentage.

```{r}
transactions %>%
  group_by(store_id) %>%
  summarize(total_sales_value = sum(sales_value, na.rm = TRUE)) %>%
  arrange(-total_sales_value)
```


```{r}
transactions %>%
  mutate(pct_loyalty_disc = 1 - (loyalty_price / regular_price)) %>%
  group_by(store_id) %>%
  summarize(avg_pct_loyalty_disc = mean(pct_loyalty_disc, na.rm = TRUE)) %>%
  arrange(desc(avg_pct_loyalty_disc))
```


# Data Visualization

The following five questions are based on concepts covered in the data visualization (`ggplot2`) slides. They can be answered using the `transactions` and `products` datasets.  

---

## Exercise 5

Create a histogram of the `quantity` variable in the `transactions` data. What, if anything, do you find unusual about this visualization? 

This question is designed to strengthen your ability to use `geom_histogram()`.

#### Answer

The unusual aspect of the histogram is its extremely long tail. The distance is so far that the histogram almost appears to be a single bar. This distortion warrants further research and cleaning of the data (e.g., `filter(quantity < 50)`).

```{r Q1, message = FALSE}
ggplot(data = transactions, aes(x = quantity)) + 
  geom_histogram()
```

---

## Exercise 6

Use a line graph to plot total sales value by `date` (I have included a mutate statement that parses the YYY-MM-DD from the `transaction_timestamp` variable and names it `date`). What kind of patterns do you find?

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_line()`.

#### Answer

There is a definite weekly pattern to transactions; likely illustrating how most folks buy the majority of their groceries on weekends.  We also see a few dates with very high and low amounts; likely related to holidays.

```{r Q2}
transactions %>% 
  mutate(date = lubridate::as_date(transaction_timestamp)) %>%
  group_by(date) %>% 
  summarize(total_sales_value = sum(sales_value, na.rm = TRUE)) %>%
  ggplot(aes(x = date, y = total_sales_value)) + 
  geom_line()
```

---

## Exercise 7

Use a bar graph to compare the total sales values of national and private-label brands. 

**Hint**: Because `transactions` does not contain product metadata, run the code below to create a new dataset with additional product information in it. Use `my_transaction_data` for your answer.

```{r}
my_transaction_data <- left_join(transactions, products, by = 'product_id')
```

This question is designed to strengthen your ability to use `dplyr` verbs in combination with `geom_col()`.

#### Answer

```{r, warning=FALSE}
my_transaction_data %>%
  group_by(brand) %>%
  summarize(total_sales_value = sum(sales_value)) %>%
  ggplot(aes(x = brand, y = total_sales_value)) + 
  geom_col()
```

---

## Exercise 8

Building on the previous exercise, suppose you want to understand whether the retailer's customers' preference for national brands (compared to private-label brands) is stronger in the soft drink category than it is in the cheese category. Examine this supposition by using a stacked bar graph to compare the split between national and private-label brands for soft drinks and cheeses. 

**Hint**: Follow these three steps to create your plot: 

 - Filter `my_transaction_data` to include only transactions with `product_category` equal to "SOFT DRINKS" or "CHEESE" 
 - Calculate total sales value by `product_category` and `brand`
 - Create the bars using `geom_col` and include `fill = brand` within `aes()`

#### Answer

```{r Q4}
my_transaction_data %>%
  filter(product_category %in% c('SOFT DRINKS', 'CHEESE')) %>%
  group_by(brand, product_category) %>%
  summarize(total_sales_value = sum(sales_value)) %>%
  ggplot(aes(x = product_category, y = total_sales_value, fill = brand)) + 
  geom_col()
```

---

## Exercise 9

The code below filters `my_transaction_data` to include only peanut better, jelly, and jam transactions. Then it creates a new variable named `package_size` equal to product size in ounces. Create a bar graph with `pb_and_j_data` to visualize the distribution of the retailer's PB\&J transactions by product size. Which two product sizes are the most popular?

```{r Q5, warning = FALSE}
pb_and_j_data <- my_transaction_data %>% 
  filter(product_category == 'PNT BTR/JELLY/JAMS') %>%
  mutate(
    product_size = as.factor(as.integer(gsub('([0-9]+)([[:space:]]*OZ)', '\\1',
                                             package_size)))
  )
```

#### Answer

The most popular product size for PB&J products is 18oz. The runner-up is 32oz.

```{r Q5a}
ggplot(pb_and_j_data, aes(x = product_size)) + 
  geom_bar()
```

This result can be confirmed numerically in the data below: 

```{r Q5b}
pb_and_j_data %>% 
  count(product_size) %>% 
  arrange(-n)
```





